\chapter{Standard Library}
\label{std}
\state{NoContent}

Standard library

\section{String}
\label{std-String}

\define[Type]{String}{define-string}{A String is a sequence of characters.}

%TODO: utf8 crap %

\section{Data Structures}
\label{std-ds}
\state{NoContent}

\subsection{Array}
\label{std-Array}

An \type{Array} is a collection of elements. It has one \tref{type parameter}{type-system-type-parameters} which corresponds to the type of these elements. Arrays can be created in three ways:

\begin{enumerate}
	\item By using their constructor: \expr{new Array()}
	\item By using \tref{array declaration syntax}{expression-array-declaration}: \expr{[1, 2, 3]}
	\item By using \tref{array comprehension}{lf-array-comprehension}: \expr{[for (i in 0...10) if (i \% 2 == 0) i]}
\end{enumerate}

Arrays come with an \href{http://api.haxe.org/Array.html}{API} to cover most use-cases. Additionally they allow read and write \tref{array access}{expression-array-access}:

\haxe{assets/ArrayAccess.hx}

Since array access in Haxe is unbounded, i.e. it is guaranteed to not throw an exception, this requires further discussion:

\begin{itemize}
	\item If a read access is made on a non-existing index, a target-dependent value is returned.
	\item If a write access is made with a positive index which is out of bounds, \expr{null} (or the \tref{default value}{define-default-value} for \tref{basic types}{types-basic-types} on \tref{static targets}{define-static-target}) is inserted at all positions between the last defined index and the newly written one.
	\item If a write access is made with a negative index, the result is unspecified.
\end{itemize}

Arrays define an \tref{iterator}{lf-iterators} over their elements. This iteration is typically optimized by the compiler to use a \tref{\expr{while} loop}{expression-while} with array index:

\haxe{assets/ArrayIterator.hx}

Haxe generates this optimized \target{Javascript} output:

\begin{lstlisting}
Main.main = function() {
	var scores = [110,170,35];
	var sum = 0;
	var _g = 0;
	while(_g < scores.length) {
		var score = scores[_g];
		++_g;
		sum += score;
	}
	console.log(sum);
};
\end{lstlisting}

Haxe does not allow arrays of mixed types unless the parameter type is forced to \tref{\type{Dynamic}}{types-dynamic}:

\haxe{assets/ArrayDynamic.hx}

\trivia{Dynamic Arrays}{In Haxe 2, mixed type array declarations were allowed. In Haxe 3, arrays can have mixed types only if they are explicitly declared as \expr{Array<Dynamic>}.}


\subsection{Vector}
\label{std-vector}

A \type{Vector} is an optimized fixed-length \emph{collection} of elements. Much like \tref{Array}{std-Array}, it has one \tref{type parameter}{type-system-type-parameters} and all elements of a vector must be of the specified type, it can be \emph{iterated over} using a \tref{for loop}{expression-for} and accessed using \tref{array access syntax}{types-abstract-array-access}. However, unlike \type{Array} and \type{List}, vector length is specified on creation and cannot be changed later.

\haxe{assets/Vector.hx}

\type{haxe.ds.Vector} is implemented as an abstract type (\ref{types-abstract}) over a native array implementation for given target and can be faster for fixed-size collections, because the memory for storing its elements is pre-allocated.

\subsection{List}
\label{std-List}
A \type{List} is a \emph{collection} for storing elements.  On the surface, a list is similar to an \Fullref{std-Array}.  However, the underlying implementation is very different.  This results in several functional differences:

\begin{enumerate}
	\item A list can not be indexed using square brackets, i.e. \expr{[0]}.
	\item A list can not be initialized.
	\item There are no list comprehensions.
	\item A list can freely modify/add/remove elements while iterating over them.
\end{enumerate}

See the \href{http://api.haxe.org/List.html}{List API} for details about the list methods.  A simple example for working with lists:
\haxe{assets/ListExample.hx}

\subsection{GenericStack}
\label{std-GenericStack}
A \type{GenericStack}, like \type{Array} and \type{List} is a container for storing elements.  It has one \tref{type parameter}{type-system-type-parameters} and all elements of the stack must be of the specified type. See the \href{http://api.haxe.org/haxe/ds/GenericStack.html}{GenericStack API} for details about its methods.  Here is a small example program for initializing and working with a \type{GenericStack}.
\haxe{assets/GenericStackExample.hx}
\trivia{FastList}{In Haxe 2, the GenericStack class was known as FastList.  Since its behavior more closely resembled a typical stack, the name was changed for Haxe 3.}
The \emph{Generic} in \type{GenericStack} is literal.  It is attributed with the \expr{:generic} metadata.  Depending on the target, this can lead to improved performance on static targets.  See \Fullref{type-system-generic} for more details.



\subsection{Map}
\label{std-Map}

A \type{Map} is a container composed of \emph{key}, \emph{value} pairs.  A \type{Map} is also commonly referred to as an associative array, dictionary, or symbol table. The following code gives a short example of working with maps:

\haxe{assets/MapExample.hx}

See the \href{http://api.haxe.org/Map.html}{Map API} for details of its methods.

Under the hood, a \type{Map} is an \tref{abstract}{types-abstract} type. At compile time, it gets converted to one of several specialized types depending on the \emph{key} type:
\begin{itemize}
	\item \type{String}: \type{haxe.ds.StringMap}
	\item \type{Int}: \type{haxe.ds.IntMap}
	\item \type{EnumValue}: \type{haxe.ds.EnumValueMap}
	\item \type{\{\}}: \type{haxe.ds.ObjectMap}
\end{itemize}

The \type{Map} type does not exist at runtime and has been replaced with one of the above objects. 

Map defines \tref{array access}{types-abstract-array-access} using its key type.


\subsection{Option}
\label{std-Option}

An option is an \tref{enum}{types-enum-instance} in the Haxe Standard Library which is defined like so:

\begin{lstlisting}
enum Option<T> {
	Some(v:T);
	None;
}
\end{lstlisting}

It can be used in various situations, such as communicating whether or not a method had a valid return and if so, what value it returned:

\haxe{assets/OptionUsage.hx}



\section{Regular Expressions}
\label{std-regex}

Haxe has built-in support for \emph{regular expressions}\footnote{http://en.wikipedia.org/wiki/Regular_expression}. They can be used to verify the format of a string, transform a string or extract some regular data from a given text.

Haxe has special syntax for creating regular expressions. We can create a regular expression object by typing it between the \expr{\textasciitilde/} combination and a single \expr{/} character:

\begin{lstlisting}
var r = ~/haxe/i;
\end{lstlisting}

Alternatively, we can create regular expression with regular syntax:

\begin{lstlisting}
var r = new EReg("haxe", "i");
\end{lstlisting}

First argument is a string with regular expression pattern, second one is a string with \emph{flags} (see below).

We can use standard regular expression patterns such as:
\begin{itemize}
	\item \expr{.} any character
	\item \expr{*} repeat zero-or-more
	\item \expr{+} repeat one-or-more
	\item \expr{?} optional zero-or-one
	\item \expr{[A-Z0-9]} character ranges
	\item \expr{[\textasciicircum\textbackslash r\textbackslash n\textbackslash t]} character not-in-range
	\item \expr{(...)} parenthesis to match groups of characters
	\item \expr{\textasciicircum} beginning of the string (beginning of a line in multiline matching mode)
	\item \expr{\$} end of the string (end of a line in multiline matching mode)
	\item \expr{|} "OR" statement.
\end{itemize}

For example, the following regular expression matches valid email addresses:
\begin{lstlisting}
~/[A-Z0-9._\%-]+@[A-Z0-9.-]+\.[A-Z][A-Z][A-Z]?/i;
\end{lstlisting}

Please notice that the \expr{i} at the end of the regular expression is a \emph{flag} that enables case-insensitive matching.

The possible flags are the following:
\begin{itemize}
	\item \expr{i} case insensitive matching
	\item \expr{g} global replace or split, see below
	\item \expr{m} multiline matching, \expr{\textasciicircum} and \expr{\$} represent the beginning and end of a line
	\item \expr{s} the dot \expr{.} will also match newlines \emph{(Neko, C++, PHP and Java targets only)}
	\item \expr{u} use UTF-8 matching \emph{(Neko and C++ targets only)}
\end{itemize}

\subsection{Matching}
\label{std-regex-match}

Probably one of the most common uses for regular expressions is checking whether a string matches the specific pattern. The \expr{match} method of a regular expression object can be used to do that:
\haxe{assets/ERegMatch.hx}

\subsection{Groups}
\label{std-regex-groups}

Specific information can be extracted from a matched string by using \emph{groups}. If \expr{match()} returns true, we can get groups using the \expr{matched(X)} method, where X is the number of a group defined by regular expression pattern:

\haxe{assets/ERegGroups.hx}

Note that group numbers start with 1 and \expr{r.matched(0)} will always return the whole matched substring.

The \expr{r.matchedPos()} will return the position of this substring in the original string:

\haxe{assets/ERegMatchPos.hx}

Additionally, \expr{r.matchedLeft()} and \expr{r.matchedRight()} can be used to get substrings to the left and to the right of the matched substring:

\haxe{assets/ERegMatchLeftRight.hx}

\subsection{Replace}
\label{std-regex-replace}

A regular expression can also be used to replace a part of the string:

\haxe{assets/ERegReplace.hx}

We can use \expr{\$X} to reuse a matched group in the replacement:

\haxe{assets/ERegReplaceGroups.hx}

\subsection{Split}
\label{std-regex-split}

A regular expression can also be used to split a string into several substrings:

\haxe{assets/ERegSplit.hx}

\subsection{Map}
\label{std-regex-map}

The \expr{map} method of a regular expression object can be used to replace matched substrings using a custom function. This function takes a regular expression object as its first argument so we may use it to get additional information about the match being done and do conditional replacement. For example:

\haxe{assets/ERegMap.hx}


\subsection{Implementation Details}
\label{std-regex-implementation-details}

Regular Expressions are implemented:

\begin{itemize}
	\item in JavaScript, the runtime is providing the implementation with the object RegExp.
	\item in Neko and C++, the PCRE library is used
	\item in Flash, PHP, C\# and Java, native implementations are used
	\item in Flash 6/8, the implementation is not available
\end{itemize}


\section{Math}
\label{std-math}

Haxe includes a floating point math library for some common mathematical operations. Most of the functions operate on and return \type{floats}. However, an \type{Int} can be used where a \type{Float} is expected, and Haxe also converts \type{Int} to \type{Float} during most numeric operations  (see \Fullref{types-numeric-operators} for more details).

Here are some example uses of the math library.  See the \href{http://api.haxe.org/Math.html}{Math API} for all available functions.

\haxe{assets/MathExample.hx}

\subsection{Special Numbers}
\label{std-math-special-numbers}

The math library has definitions for several special numbers:

\begin{itemize}
	\item NaN (Not a Number): returned when a mathmatically incorrect operation is executed, e.g. Math.sqrt(-1)
	\item POSITIVE_INFINITY: e.g. divide a positive number by zero
	\item NEGATIVE_INFINITY: e.g. divide a negative number by zero
	\item PI : 3.1415...
\end{itemize}

\subsection{Mathematical Errors}
\label{std-math-mathematical-errors}
Although neko can fluidly handle mathematical errors, like division by zero, this is not true for all targets.  Depending on the target, mathematical errors may produce exceptions and ultimately errors.

\subsection{Integer Math}
\label{std-math-integer-math}

If you are targeting a platform that can utilize integer operations, e.g. integer division, it should be wrapped in \emph{Std.int()} for improved performance.  The Haxe Compiler can then optimize for integer operations.  An example:

\begin{lstlisting}
	var intDivision = Std.int(6.2/4.7);
\end{lstlisting}

\todo{I think C++ can use integer operatins, but I don't know about any other targets. Only saw this mentioned in an old discussion thread, still true?}

\subsection{Extensions}
\label{std-math-extensions}
It is common to see \Fullref{lf-static-extension} used with the math library.  This code shows a simple example:  
\haxe{assets/MathStaticExtension.hx}
\haxe{assets/MathExtensionUsage.hx}


\section{Lambda}
\label{std-Lambda}

\define{Lambda}{define-lambda}{Lambda is a functional language concept within Haxe that allows you to apply a function to a list or \tref{iterators}{lf-iterators}. The Lambda class is a collection of functional methods in order to use functional-style programming with Haxe.}

It is ideally used with \expr{using Lambda} (see \tref{Static Extension}{lf-static-extension}) and then acts as an extension to \type{Iterable} types. 

On static platforms, working with the \type{Iterable} structure might be slower than performing the operations directly on known types, such as \type{Array} and \type{List}.

\paragraph{Lambda Functions}
The Lambda class allows us to operate on an entire \type{Iterable} at once.
This is often preferable to looping routines since it is less error prone and easier to read. 
For convenience, the \type{Array} and \type{List} class contains some of the frequently used methods from the Lambda class.

It is helpful to look at an example. The exists function is specified as:

\begin{lstlisting}
static function exists<A>( it : Iterable<A>, f : A -> Bool ) : Bool
\end{lstlisting}

Most Lambda functions are called in similar ways. The first argument for all of the Lambda functions is the \type{Iterable} on which to operate. Many also take a function as an argument.

\begin{description}
	\item[\expr{Lambda.array}, \expr{Lambda.list}] Convert Iterable to \type{Array} or \type{List}. It always returns a new instance.
	\item[\expr{Lambda.count}] Count the number of elements.  If the Iterable is a \type{Array} or \type{List} it is faster to use its length property.
	\item[\expr{Lambda.empty}] Determine if the Iterable is empty. For all Iterables it is best to use the this function; it's also faster than compare the length (or result of Lambda.count) to zero.
	\item[\expr{Lambda.has}] Determine if the specified element is in the Iterable.
	\item[\expr{Lambda.exists}] Determine if criteria is satisfied by an element.
	\item[\expr{Lambda.indexOf}] Find out the index of the specified element.
	\item[\expr{Lambda.find}] Find first element of given search function.
	\item[\expr{Lambda.foreach}] Determine if every element satisfies a criteria.
	\item[\expr{Lambda.iter}] Call a function for each element.
	\item[\expr{Lambda.concat}] Merge two Iterables, returning a new List.
	\item[\expr{Lambda.filter}] Find the elements that satisfy a criteria, returning a new List.
	\item[\expr{Lambda.map}, \expr{Lambda.mapi}] Apply a conversion to each element, returning a new List.
	\item[\expr{Lambda.fold}] Functional fold, which is also known as reduce, accumulate, compress or inject.
\end{description}

This example demonstrates the Lambda filter and map on a set of strings:

\begin{lstlisting}
using Lambda;
class Main {
    static function main() {
        var words = ['car', 'boat', 'cat', 'frog'];

		var isThreeLetters = function(word) return word.length == 3;
		var capitalize = function(word) return word.toUpperCase();
		
		// Three letter words and capitalized. 
		trace(words.filter(isThreeLetters).map(capitalize)); // [CAR,CAT]
    }
}
\end{lstlisting} 

This example demonstrates the Lambda count, has, foreach and fold function on a set of ints.

\begin{lstlisting}
using Lambda;
class Main {
    static function main() {
        var numbers = [1, 3, 5, 6, 7, 8];
		
		trace(numbers.count()); // 6
		trace(numbers.has(4)); // false
		
        // test if all numbers are greater/smaller than 20
		trace(numbers.foreach(function(v) return v < 20)); // true
        trace(numbers.foreach(function(v) return v > 20)); // false
		
        // sum all the numbers
		var sum = function(num, total) return total += num;
		trace(numbers.fold(sum, 0)); // 30
    }
}
\end{lstlisting} 

\section{Template}
\label{std-template}

Haxe comes with a standard template system with an easy to use syntax which is interpreted by a lightweight class called \type{haxe.Template}.

A template is a string or a file that is used to produce any kind of string output depending on the input. Here is a small template example:

\haxe{assets/Template.hx}

The console will trace \ic{My name is Mark, 30 years old}.

\paragraph{Expressions}
An expression can be put between the \ic{::}, the syntax allows the current possibilities:

\begin{description}
	\item[\ic{::name::}] the variable name
	\item[\ic{::expr.field::}] field access
	\item[\ic{::(expr)::}] the expression expr is evaluated
	\item[\ic{::(e1 op e2)::}] the operation op is applied to e1 and e2
	\item[\ic{::(135)::}] the integer 135. Float constants are not allowed
\end{description}

\paragraph{Conditions}
It is possible to test conditions using \ic{::if flag1::}. Optionally, the condition may be followed by \ic{::elseif flag2::} or \ic{::else::}. Close the condition with \ic{::end::}.

\begin{lstlisting} 
::if isValid:: valid ::else:: invalid ::end::
\end{lstlisting} 

Operators can be used but they don't deal with operator precedence. Therefore it is required to enclose each operation in parentheses \ic{()}. Currently, the following operators are allowed: \ic{+}, \ic{-}, \ic{*}, \ic{/}, \ic{>}, \ic{<},  \ic{>=}, \ic{<=}, \ic{==}, \ic{!=}, \ic{&&} and \ic{||}.

For example \ic{::((1 + 3) == (2 + 2))::} will display true. 

\begin{lstlisting} 
::if (points == 10):: Great! ::end::
\end{lstlisting} 

To compare to a string, use double quotes \ic{"} in the template.
\begin{lstlisting} 
::if (name == "Mark"):: Hi Mark ::end::
\end{lstlisting} 

\paragraph{Iterating}
Iterate on a structure by using \ic{::foreach::}. End the loop with \ic{::end::}.
\begin{lstlisting} 
<table>
	<tr>
		<th>Name</th>
		<th>Age</th>
	</tr>
	::foreach users::
		<tr>
			<td>::name::</td>
			<td>::age::</td>
		</tr>
	::end::
</table>
\end{lstlisting} 

\paragraph{Sub-templates}
To include templates in other templates, pass the sub-template result string as a parameter.
\begin{lstlisting} 
var users = [{name:"Mark", age:30}, {name:"John", age:45}];

var userTemplate = new haxe.Template("::foreach users:: ::name::(::age::) ::end::");
var userOutput = userTemplate.execute({users: users});

var template = new haxe.Template("The users are ::users::");
var output = template.execute({users: userOutput});
trace(output);
\end{lstlisting} 
The console will trace \ic{The users are Mark(30) John(45)}.

\paragraph{Template macros}
To call custom functions while parts of the template are being rendered, provide a \expr{macros} object to the argument of \expr{Template.execute}. The key will act as the template variable name, the value refers to a callback function that should return a \type{String}. The first argument of this macro function is always a \expr{resolve()} method, followed by the given arguments. The resolve function can be called to retrieve values from the template context. If \expr{macros} has no such field, the result is unspecified.

The following example passes itself as macro function context and executes \ic{display} from the template.
\haxe{assets/TemplateMacros.hx}
The console will trace \ic{The results: Mark ran 3.5 kilometers in 15 minutes}.

\paragraph{Globals}
Use the \expr{Template.globals} object to store values that should be applied across all \type{haxe.Template} instances. This has lower priority than the context argument of \expr{Template.execute}.

\paragraph{Using resources}

To separate the content from the code, consider using the \tref{resource embedding system}{cr-resources}. 
Place the template-content in a new file called \ic{sample.mtt}, add \ic{-resource sample.mtt@my_sample} to the compiler arguments and retrieve the content using \expr{haxe.Resource.getString}.
\haxe{assets/TemplateResource.hx}

When running the template system on the server side, you can simply use \expr{neko.Lib.print} or \expr{php.Lib.print} instead of trace to display the HTML template to the user.


\section{Reflection}
\label{std-reflection}

Haxe supports runtime reflection of types and fields. Special care has to be taken here because runtime representation generally varies between targets. In order to use reflection correctly it is necessary to understand what kind of operations are supported and what is not. Given the dynamic nature of reflection, this can not always be determined at compile-time.

The reflection API consists of two classes:

\begin{description}
	\item[Reflect:] A lightweight API which work best on \tref{anonymous structures}{types-anonymous-structure}, with limited support for \tref{classes}{types-class-instance}. 
	\item[Type:] A more robust API for working with classes and \tref{enums}{types-enum-instance}.
\end{description}

The available methods are detailed in the API for \href{http://api.haxe.org//Reflect.html}{Reflect} and \href{http://api.haxe.org//Type.html}{Type}.

Reflection can be a powerful tool, but it is important to understand why it can also cause problems. As an example, several functions expect a \tref{String}{std-String} argument and try to resolve it to a type or field. This is vulnerable to typing errors:

\haxe{assets/ReflectionTypo.hx}

However, even if there are no typing errors it is easy to come across unexpected behavior:

\haxe{assets/ReflectionMissingType.hx}

The problem here is that the compiler never actually ``sees'' the type \type{haxe.Template}, so it does not compile it into the output. Furthermore, even if it were to see the type there could be issues arising from \tref{dead code elimitation}{cr-dce} eliminating types or fields which are only used via reflection.

Another set of problems comes from the fact that, by design, several reflection functions expect arguments of type \tref{Dynamic}{types-dynamic}, meaning the compiler cannot check if the passed in arguments are correct. The following example demonstrates a common mistake when working with \expr{callMethod}:

\haxe{assets/ReflectionWrongUsage.hx}

The commented out call would be accepted by the compiler because it assigns the string \expr{"f"} to the function argument \expr{func} which is specified to be \expr{Dynamic}.

A good advice when working with reflection is to wrap it in a few functions within an application or API which are called by otherwise type-safe code. An example could look like this:

\haxe{assets/ReflectionWrap.hx}

While the method \expr{reflective} could interally work with reflection (and \type{Dynamic} for that matter) a lot, its return value is a typed structure which the callers can use in a type-safe manner.


\section{Serialization}
\label{std-serialization}

Many runtime values can be serialized and deserialized using the \type{haxe.Serializer} and \type{haxe.Unserializer} classes. Both support two usages:

\begin{enumerate}
	\item Create an instance and continuously call the \expr{serialize}/\expr{unserialize} method to handle multiple values.
	\item Call their static \expr{run} method to serialize/deserialize a single value.
\end{enumerate}

The following example demonstrates the first usage:

\haxe{assets/SerializationExample.hx}

The result of the serialization (here stored in local variable \expr{s}) is a \tref{String}{std-String} and can be passed around at will, even remotely. Its format is described in \Fullref{std-serialization-format}.

\paragraph{Supported values}

\begin{itemize}
	\item \expr{null}
	\item \type{Bool}, \type{Int} and \type{Float} (including infinities and \expr{NaN})
	\item \type{String}
	\item \type{Date}
	\item \type{haxe.io.Bytes} (encoded as base64)
	\item \tref{\type{Array}}{std-Array} and \tref{\type{List}}{std-List}
	\item \type{haxe.ds.StringMap}, \type{haxe.ds.IntMap} and \type{haxe.ds.ObjectMap}
	\item \tref{anonymous structures}{types-anonymous-structure}
	\item Haxe \tref{class instances}{types-class-instance} (not native ones)
	\item \tref{enum instances}{types-enum-instance}
\end{itemize}

\paragraph{Serialization configuration}

Serialization can be configured in two ways. For both a static variable can be set to influence all \type{haxe.Serializer} instances, and a member variable can be set to only influence a specific instance:

\begin{description}
	\item[\expr{USE_CACHE}, \expr{useCache}:] If true, repeated structures or class\slash enum instances are serialized by reference. This can avoid infinite loops for recursive data at the expense of longer serialization time. By default, object caching is disabled; strings however are always cached.
	\item[\expr{USE_ENUM_INDEX}, \expr{useEnumIndex}:] If true, enum constructors are serialized by their index instead of their name. This can make the resulting string shorter, but breaks if enum constructors are inserted into the type before deserialization. This behavior is disabled by default.
\end{description}

\paragraph{Deserialization behavior}

If the serialization result is stored and later used for deserialization, care has to be taken to maintain compatibility when working with class and enum instances. It is then important to understand exactly how unserialization is implemented.

\begin{itemize}
	\item The type has to be available in the runtime where the deserialization is made. If \tref{dead code elimination}{cr-dce} is active, a type which is used only through serialization might be removed.
	\item Each \type{Unserializer} has a member variable \expr{resolver} which is used to resolve classes and enums by name. Upon creation of the \type{Unserializer} this is set to \expr{Unserializer.DEFAULT_RESOLVER}. Both that and the instance member can be set to a custom resolver.
	\item Classes are resolved by name using \expr{resolver.resolveClass(name)}. The instance is then created using \expr{Type.createEmptyInstance}, which means that the class constructor is not called. Finally, the instance fields are set according to the serialized value.
	\item Enums are resolved by name using \expr{resolver.resolveEnum(name)}. The enum instance is then created using \expr{Type.createEnum}, using the serialized argument values if available. If the constructor arguments were changed since serialization, the result is unspecified.
\end{itemize}

\paragraph{Custom (de)serialization}

If a class defines the member method \expr{hxSerialize}, that method is called by the serializer and allows custom serialization of the class. Likewise, if a class defines the member method \expr{hxUnserialize} it is called by the deserializer:

\haxe{assets/SerializationCustom.hx}

In this example we decide that we want to ignore the value of member variable \expr{y} and do not serialize it. Instead we default it to \expr{-1} in \expr{hxUnserialize}. Both methods are annotated with the \expr{:keep} metadata to prevent \tref{dead code elimination}{cr-dce} from removing them as they are never properly referenced in the code.


\subsection{Serialization format}
\label{std-serialization-format}

Each supported value is translated to a distinct prefix character, followed by the necessary data.

\begin{description}
	\item[\expr{null}:] \expr{n}
	\item[\type{Int}:] \expr{z} for zero, or \expr{i} followed by the integer display (e.g. \expr{i456})
	\item[\type{Float}:] \mbox{}
		\begin{description}
			\item[\expr{NaN}:] \expr{k}
			\item[negative infinity:] \expr{m}
			\item[positive infinity:] \expr{p}
			\item[finite floats:] \expr{d} followed by the float display (e.g. \expr{d1.45e-8})
		\end{description}
	\item[\type{Bool}:] \expr{t} for \expr{true}, \expr{f} for \expr{false}
	\item[\type{String}:] \expr{y} followed by the url encoded string length, then \expr{:} and the url encoded string (e.g. \expr{y10:hi\%20there for "hi there".}
	\item[name-value pairs:] a serialized string representing the name followed by the serialized value
	\item[structure:] \expr{o} followed by the list of name-value pairs and terminated by \expr{g} (e.g. \expr{oy1:xi2y1:kng} for \expr{\{x:2, k:null\}})
	\item[\type{List}:] \expr{l} followed by the list of serialized items, followed by \expr{h} (e.g. \expr{lnnh} for a list of two \expr{null} values)
	\item[\type{Array}:] \expr{a} followed by the list of serialized items, followed by \expr{h}. For multiple consecutive \expr{null} values, \expr{u} followed by the number of \expr{null} values is used (e.g. \expr{ai1i2u4i7ni9h for [1,2,null,null,null,null,7,null,9]})
	\item[\type{Date}:] \expr{v} followed by the date itself (e.g. \expr{v2010-01-01 12:45:10})
	\item[\type{haxe.ds.StringMap}:] \expr{b} followed by the name-value pairs, followed by \expr{h} (e.g. \expr{by1:xi2y1:knh} for \expr{\{"x" => 2, "k" => null\}})
	\item[\type{haxe.ds.IntMap}:] \expr{q} followed by the key-value pairs, followed by \expr{h}. Each key is represented as \expr{:<int>} (e.g. \expr{q:4n:5i45:6i7h} for \expr{\{4 => null, 5 => 45, 6 => 7\}})
	\item[\type{haxe.ds.ObjectMap}:] \expr{M} followed by serialized value pairs representing the key and value, followed by \expr{h}
	\item[\type{haxe.io.Bytes}:] \expr{s} followed by the length of the base64 encoded bytes, then \expr{:} and the byte representation using the codes \expr{A-Za-z0-9\%} (e.g. \expr{s3:AAA} for 2 bytes equal to \expr{0}, and \expr{s10:SGVsbG8gIQ} for \expr{haxe.io.Bytes.ofString("Hello !")})
	\item[exception:] \expr{x} followed by the exception value
	\item[class instance:] \expr{c} followed by the serialized class name, followed by the name-value pairs of the fields, followed by \expr{g} (e.g. \expr{cy5:Pointy1:xzy1:yzg} for \expr{new Point(0, 0)} (having two integer fields \expr{x} and \expr{y})
        \item[enum instance (by name):] \expr{w} followed by the serialized enum name, followed by the serialized constructor name, followed by \expr{:}, followed by the number of arguments, followed by the argument values (e.g. \expr{wy3:Fooy1:A:0} for \expr{Foo.A} (with no arguments), \expr{wy3:Fooy1:B:2i4n} for \expr{Foo.B(4,null)})
	\item[enum instance (by index):] \expr{j} followed by the serialized enum name, followed by \expr{:}, followed by the constructor index (starting from 0), followed by \expr{:}, followed by the number of arguments, followed by the argument values (e.g. \expr{wy3:Foo:0:0} for \expr{Foo.A} (with no arguments), \expr{wy3:Foo:1:2i4n} for \expr{Foo.B(4,null)})
	\item[cache references:] \mbox{}
		\begin{description}
			\item[\type{String}:] \expr{R} followed by the corresponding index in the string cache (e.g. \expr{R456})
			\item[class, enum or structure] \expr{r} followed by the corresponding index in the object cache (e.g. \expr{r42})
		\end{description}
	\item[custom:] \expr{C} followed by the class name, followed by the custom serialized data, followed by \expr{g}
\end{description}

\noindent Cached elements and enum constructors are indexed from zero.

\section{Json}
\label{std-Json}

Haxe provides built-in support for (de-)serializing \emph{JSON}\footnote{http://en.wikipedia.org/wiki/JSON} data via the \type{haxe.Json} class.

\subsection{Parsing JSON}
\label{std-Json-parsing}

Use the \expr{haxe.Json.parse} static method to parse \emph{JSON} data and obtain a Haxe value from it:
\haxe{assets/JsonParse.hx}

Note that the type of the object returned by \expr{haxe.Json.parse} is \expr{Dynamic}, so if the structure of our data is well-known, we may want to specify a type using \tref{anonymous structures}{types-anonymous-structure}. This way we provide compile-time checks for accessing our data and most likely more optimal code generation, because compiler knows about types in a structure:
\haxe{assets/JsonParseTyped.hx}

\subsection{Encoding JSON}
\label{std-Json-encoding}

Use the \expr{haxe.Json.stringify} static method to encode a Haxe value into a \emph{JSON} string:
\haxe{assets/JsonStringify.hx}

\subsection{Implementation details}
\label{std-Json-implementation-details}

The \type{haxe.Json} API automatically uses native implementation on targets where it is available, i.e. \emph{JavaScript}, \emph{Flash} and \emph{PHP} and provides its own implementation for other targets.

Usage of Haxe own implementation can be forced with \expr{-D haxeJSON} compiler argument. This will also provide serialization of \tref{enums}{types-enum-instance} by their index, \tref{maps}{std-Map} with string keys and class instances.

Older browsers (Internet Explorer 7, for instance) may not have native \emph{JSON} implementation. In case it's required to support them, we can include one of the JSON implementations available on the internet in the HTML page. Alternatively, a \expr{-D old_browser} compiler argument that will make \type{haxe.Json} try to use native JSON and, in case it's not available, fallback to its own implementation.

\section{Xml}
\label{std-Xml}

\section{Input/Output}
\label{std-input-output}

\section{Sys/sys}
\label{std-sys}

\section{Remoting}
\label{std-remoting}

Haxe remoting is a way to communicate between different platforms. With Haxe remoting, applications can transmit data transparently, send data and call methods between server and client side.

\subsection{Remoting Connection}
\label{std-remoting-connection}

In order to use remoting, there must be a connection established. There are two kinds of Haxe Remoting connections: 
\begin{description}
	\item[\expr{haxe.remoting.Connection}] is used for \emph{synchronous connections}, where the results can be directly obtained when calling a method. 
	\item[\expr{haxe.remoting.AsyncConnection}] is used for \emph{asynchronous connections}, where the results are events that will happen later in the execution process.
\end{description}

\paragraph{Start a connection}
There are some target-specific constructors with different purposes that can be used to set up a connection:

\begin{description}
	\item[All targets:]
		\begin{description}
			\item[\expr{HttpAsyncConnection.urlConnect(url:String)}]  
				Returns an asynchronous connection to the given URL which should link to a Haxe server application. 
		\end{description}
		
	\item[Flash:]
		\begin{description}
			\item[\expr{ExternalConnection.jsConnect(name:String, ctx:Context)}]  
				Allows a connection to the local JavaScript Haxe code. The JS Haxe code must be compiled with the class ExternalConnection included. This only works with Flash Player 8 and higher.
			\item[\expr{AMFConnection.urlConnect(url:String)} and \expr{AMFConnection.connect( cnx : NetConnection )}]  
				Allows a connection to an \href{http://en.wikipedia.org/wiki/Action_Message_Format}{AMF Remoting server} such as \href{http://www.adobe.com/products/adobe-media-server-family.html}{Flash Media Server} or \href{http://www.silexlabs.org/amfphp/}{AMFPHP}.
			\item[\expr{SocketConnection.create(sock:flash.XMLSocket)}]  
				Allows remoting communications over an \type{XMLSocket}
			\item[\expr{LocalConnection.connect(name:String)}]  
				Allows remoting communications over a \href{http://api.haxe.org/haxe/remoting/LocalConnection.html}{Flash LocalConnection}
		\end{description}
		
	\item[Javascript:]
		\begin{description}
			\item[\expr{ExternalConnection.flashConnect(name:String, obj:String, ctx:Context)}]  
				Allows a connection to a given Flash Object. The Haxe Flash content must be loaded and it must include the \expr{haxe.remoting.Connection} class. This only works with Flash 8 and higher. 
		\end{description}
		
	\item[Neko:]
		\begin{description}
			\item[\expr{HttpConnection.urlConnect(url:String)}]  
				Will work like the asynchronous version but in synchronous mode.
			\item[\expr{SocketConnection.create(...)}]  
				Allows real-time communications with a Flash client which is using an \type{XMLSocket} to connect to the server.
		\end{description}
\end{description}

\paragraph{Remoting context}

Before communicating between platforms, a remoting context has to be defined. This is a shared API that can be called on the connection at the client code.

This server code example creates and shares an API:
\begin{lstlisting}
class Server {
	function new() { }
	function foo(x, y) { return x + y; }

	static function main() {
		var ctx = new haxe.remoting.Context();
		ctx.addObject("Server", new Server());
		
		if(haxe.remoting.HttpConnection.handleRequest(ctx))
		{
			return;
		}
		
		// handle normal request
		trace("This is a remoting server !");
	} 
}
\end{lstlisting}

\paragraph{Using the connection}

Using a connection is pretty convenient. Once the connection is obtained, use classic dot-access to evaluate a path and then use \expr{call()} to call the method in the remoting context and get the result.
The asynchronous connection takes an additional function parameter that will be called when the result is available.

This client code example connects to the server remoting context and calls a function \expr{foo()} on its API.
\begin{lstlisting}
class Client {
  static function main() {
    var cnx = haxe.remoting.HttpAsyncConnection.urlConnect("http://localhost/");
    cnx.setErrorHandler( function(err) trace('Error: $err'); } );
    cnx.Server.foo.call([1,2], function(data) trace('Result: $data'););
  }
}
\end{lstlisting}

To make this work for the Neko target, setup a Neko Web Server, point the url in the Client to \ic{"http://localhost2000/remoting.n"} and compile the Server using \ic{-main Server -neko remoting.n}.

\paragraph{Error handling}

\begin{itemize}
	\item When an error occurs in a asynchronous call, the error handler is called as seen in the example above.
	\item When an error occurs in a synchronous call, an exception is raised on the caller-side as if we were calling a local method.
\end{itemize}

\paragraph{Data serialization}

Haxe Remoting can send a lot of different kinds of data. See \tref{Serialization}{std-serialization}.

\subsection{Implementation details}
\label{std-remoting-implementation-details}

\paragraph{Javascript security specifics}

The html-page wrapping the js client must be served from the same domain as the one where the server is running. The same-origin policy restricts how a document or script loaded from one origin can interact with a resource from another origin. The same-origin policy is used as a means to prevent some of the cross-site request forgery attacks.

To use the remoting across domain boundaries, CORS (cross-origin resource sharing) needs to be enabled by defining the header \ic{X-Haxe-Remoting} in the \ic{.htaccess}:

\begin{lstlisting} 
# Enable CORS
Header set Access-Control-Allow-Origin "*"
Header set Access-Control-Allow-Methods: "GET,POST,OPTIONS,DELETE,PUT"
Header set Access-Control-Allow-Headers: X-Haxe-Remoting
\end{lstlisting} 

See \href{http://en.wikipedia.org/wiki/Same-origin_policy}{same-origin policy} for more information on this topic.

Also note that this means that the page can't be served directly from the file system \ic{"file:///C:/example/path/index.html"}.

\paragraph{Flash security specifics}

When Flash accesses a server from a different domain, set up a \ic{crossdomain.xml} file on the server, enabling the \ic{X-Haxe} headers.

\begin{lstlisting} 
<cross-domain-policy>
	<allow-access-from domain="*"/> <!-- or the appropriate domains -->
	<allow-http-request-headers-from domain="*" headers="X-Haxe*"/>
</cross-domain-policy>
\end{lstlisting} 

\paragraph{Arguments types are not ensured}

There is no guarantee of any kind that the arguments types will be respected when a method is called using remoting. 
That means even if the arguments of function \expr{foo} are typed to \type{Int}, the client will still be able to use strings while calling the method. 
This can lead to security issues in some cases. When in doubt, check the argument type when the function is called by using the \expr{Std.is} method.


\section{SPOD}
\label{std-spod}
\flag{fold}{true}

The Simple Persistent Objects Database library, also known as SPOD, makes it possible to write Haxe expressions instead of database queries, while keeping type-safety.

\subsection{SPOD Objects}
\label{std-spod-objects}

Declare a SPOD Object by extending the \expr{sys.db.Object} class:

\begin{lstlisting} 
import sys.db.Types;
class User extends sys.db.Object {
    public var id:SId;
    public var name:SString<32>;
    public var birthday:SDate;
    public var phoneNumber:SNull<SText>;
}
\end{lstlisting} 

Special types are declared in \expr{sys.db.Types} in order to provide additional information for SPOD. See \tref{complete list of types and related SQL types}{std-spod-supported-types}.

\paragraph{Insert}
In order to insert a new SPOD, simply do the following:

\begin{lstlisting} 
var user = new User();
user.name = "Random156";
user.birthday = Date.now();
user.insert();
\end{lstlisting} 

After the \expr{insert()} is done, the auto increment unique id will be set and all fields that were null but not declared as nullable will be set to their default value (\expr{0} for numbers, \expr{""} for strings and empty bytes for binaries).

\paragraph{Update}
Once there is an instance of the SPOD object, modify its fields and call \expr{update()} to send these changes to the database.

\begin{lstlisting} 
var user = User.manager.get(1);
if( user.phoneNumber == null) user.phoneNumber = "+3360000000";
user.update();
\end{lstlisting} 

\paragraph{Delete}
Use \expr{delete()} to delete the SPOD object from the database.

\begin{lstlisting} 
var user = User.manager.get(1);
if (user != null) user.delete();
\end{lstlisting} 

\subsection{Initialization}
\label{std-spod-initialization}

There are two static methods that might be needed to call before/after using SPOD:

\begin{description}
	\item[\expr{sys.db.Manager.initialize()}] will initialize the created managers. Make sure to call it at least once before using SPOD.
	\item[\expr{sys.db.Manager.cleanup()}] will cleanup the temporary object cache. This can be done if you are using server module caching to free memory or after a rollback to make sure that we don't use the cached object version.
\end{description}

\subsection{Making connection}
\label{std-spod-connection}

Connect to your database and create tables directly from code without writing SQL. For instance, by using \type{sys.db.Mysql}. 

\begin{lstlisting} 
var cnx = sys.db.Mysql.connect({
   host : "localhost",
   port : null,
   user : "root",
   pass : "",
   database : "testBase",
   socket : null,
});
sys.db.Manager.cnx = cnx;
\end{lstlisting} 

\subsection{Creating tables}
\label{std-spod-tables}

Calling \expr{sys.db.TableCreate.create} will execute a \ic{CREATE TABLE SQL} request based on the SPOD infos.

\begin{lstlisting} 
if (!sys.db.TableCreate.exists(User.manager)) {
    sys.db.TableCreate.create(User.manager);
}
\end{lstlisting} 

Please note that currently \type{TableCreate} will not create the index or initialize the relations of your table.
\todo{Find out if this is still true and if this is something that can be fixed}

\subsection{SPOD Supported types}
\label{std-spod-supported-types}

\begin{center}
\begin{tabular}{| l | l | l |}
	\hline
	\multicolumn{3}{|c|}{Supported types} \\ \hline
	Haxe Type &  SQL Type  &  Description \\ \hline
	\type{Int}, \expr{SInt} & \expr{SSQL} INT & a classic 32 bits signed integer \\
	\type{Float}, \expr{SFloat} & \expr{SQL DOUBLE} & a double precision float value  \\
	\type{Bool} \expr{SBool}  & \expr{SQL TINYINT(1)} or \expr{BOOL} & a boolean value \\
	\type{Date}, \expr{SDateTime}  & \expr{SQL DATETIME} & a complete date value \\
	\type{SDate}  & \expr{SQL DATE} & a date-only value \\
	\type{SString<K>}  & \expr{SQL VARCHAR(K)} & a size-limited string value \\
	\type{String}, \expr{SText}  & \expr{SQL MEDIUMTEXT} & a text up to 16 MB \\
	\type{SBytes<K>}  & \expr{SQL BINARY(K)} & a fixed-size bytes value \\
	\type{SBinary}, \type{haxe.io.Bytes}  & \expr{SQL MEDIUMBLOB} & up to 16 MB bytes \\
	\type{SId}  & \expr{SQL INT AUTO INCREMENT} & same as \expr{SInt} but used as an unique ID with auto increment \\
	\type{SEnum<E>}  & \expr{SQL TINYINT UNSIGNED} & a single enum without parameters which index is stored as a small integer \\
	\type{SFlags<E>}  &  & a 32 bits flag that uses an enum as bit markers. See \type{haxe.EnumFlags} \\
	\type{SData<Anything>}  &   & allow arbitrary serialized data (see below) \\
	\type{Null<T>}, \type{SNull<T>} & & tells that this field can be \ic{NULL} in the database \\
\end{tabular}
\end{center}

\paragraph{Advanced Types}

The following advanced types are also available to use specific storage size:

\begin{center}
\begin{tabular}{| l | l | l |}
	\hline
	\multicolumn{3}{|c|}{Supported advanced types} \\ \hline
	Haxe Type &  SQL Type  &  Description \\ \hline
	\expr{SUInt} & \expr{SQL UNSIGNED INT} & an unsigned 32 bits integer \\
	\expr{STinyInt}, \expr{STinyUInt} & \expr{SQL TINYINT} & a small 8 bits signed/unsigned integer \\
	\expr{SSmallInt}, \expr{SSmallUInt} & \expr{SQL SMALLINT} & a small 16 bits signed/unsigned integer \\
	\expr{SMediumIInt}, \expr{SMediumUInt} & \expr{SQL MEDIUMINT} & a small 24 bits signed/unsigned integer \\
	\expr{SBigInt} & \expr{SQL BIGINT} & a 64 bits signed integer - typed as Float in Haxe \\
	\expr{SSingle} & \expr{SQL FLOAT} & a single precision float value \\
	\expr{STinyText} & \expr{SQL TINYTEXT} & a text up to 255 bytes \\
	\expr{SSmallText} & \expr{SQL TEXT} & a text up to 65KB \\
	\expr{STimeStamp} & \expr{SQL TIMESTAMP} & a 32-bits date timestamp \\
	\expr{SSmallBinary} & \expr{SQL BLOB} & up to 65 KB bytes \\
	\expr{SLongBinary} & \expr{SQL LONGBLOB} & up to 4GB bytes \\
	\expr{SUId} & \expr{SQL INT UNSIGNED AUTO INCREMENT} & same as \expr{SUInt} but used as an unique ID with auto increment \\
	\expr{SBigId} & \expr{SQL BIGINT AUTO INCREMENT} & same as \expr{SBigInt} but used as an unique ID with auto increment - compiled as \type{Float} in Haxe \\
	\expr{SSmallFlags<E>} &   & similar to \expr{SFlags} except that the integer used to store the data is based on the number of flags allowed \\
\end{tabular}
\end{center}
	
	
\subsection{SPOD Metadata}
\label{std-spod-metadata}

Metadata can be added to SPOD classes to declare additional information that will be used by SPOD.

\begin{center}
\begin{tabular}{| l | l |}
	\hline
	\multicolumn{2}{|c|}{Before each class field} \\ \hline
	Metatag &  Description \\ \hline
	\expr{@:skip} & ignore this field, which will not be part of the database schema \\
	\expr{@:relation} & declare this field as a relation (see \tref{relations}{std-spod-relations}) \\
\end{tabular}
\end{center}
	
\begin{center}
\begin{tabular}{| l | l |}
	\hline
	\multicolumn{2}{|c|}{Before the SPOD class} \\ \hline
	Metatag &  Description \\ \hline
	\expr{@:table("myTableName")} & change the table name (by default it's the same as the class name) \\
	\expr{@:id(field1,field2,...)} & specify the primary key fields for this table.  \\
	\expr{@:index(field1,field2,...,[unique])} & declare an index consisting of the specified classes fields - in that order. If the last field is unique then it means that's an unique index (each combination of fields values can only occur once) \\
\end{tabular}
\end{center}	
	
The following class does not have a unique id with auto increment, but a two-fields unique primary key:

\begin{lstlisting} 
@:id(uid, gid)
class UserGroup extends sys.db.Object {
    public var uid:SInt;
    public var gid:SInt;
}
\end{lstlisting} 

\subsection{SPOD Manager}
\label{std-spod-manager}

\paragraph{Manager}

Each SPOD object need its own manager. The SPOD Macros will add this manager automatically. To define a custom manager (should extend \type{sys.db.Manager}), add the following line to your SPOD class body:

\begin{lstlisting} 
public static var manager = new sys.db.Manager<User>(User);
\end{lstlisting} 


\paragraph{Get}

In order to retrieve an instance of SPOD, call the \expr{Manager.get} method by using the object unique identifier (primary key).

\begin{lstlisting} 
var user = User.manager.get(1);
if (user == null) throw "User #1 not found";
trace(u.name);
\end{lstlisting} 

If a primary key with multiple values is defined, use the following declaration:

\begin{lstlisting} 
var userGroup = UserGroup.manager.get({uid:1, gid:2});
\end{lstlisting} 


\paragraph{Select}
To return the first result object, use the \expr{Manager.select} method.

\begin{lstlisting} 
var user = User.manager.select($name == "John");
\end{lstlisting} 

\paragraph{Count}
Use the \expr{Manager.count} method to count the number of objects matching the given search query:

\begin{lstlisting} 
var count = User.manager.count($name.like("J\%") && $phoneNumber != null);
\end{lstlisting} 

\paragraph{Delete}
Delete all objects matching the given query:

\begin{lstlisting} 
User.manager.delete($id > 1000);
\end{lstlisting} 


\paragraph{Search}

To search for objects, use the \expr{Manager.search} method.

\begin{lstlisting} 
var minId = 10;
for(user in User.manager.search($id < minId)) {
    trace(user);
}
\end{lstlisting} 

In order to differentiate between the database fields and the Haxe variables, all the database fields are prefixed with a dollar in search queries.

Search queries are checked at compile-time and the following SQL code is generated instead:

\begin{lstlisting} 
unsafeSearch("SELECT * FROM User WHERE id < " + Manager.quoteInt(minId));
\end{lstlisting} 

The code generator also makes sure that no SQL injection is ever possible.

\paragraph{Supported search syntax}

\begin{description}
	\item[constants] integers, floats, strings, null, true, false
	\item[all operations] \expr{+}, \expr{-}, \expr{*}, \expr{/}, \expr{\%}, \expr{|}, \expr{&}, \expr{^}, \expr{>>}, \expr{<<}, \expr{>>>}
	\item[unary operations] \expr{!}, \expr{-}, \expr{~}
	\item[all comparisons] \expr{==} , \expr{>=} , \expr{<=}, \expr{>}, \expr{<}, \expr{!=}
	\item[bool tests] \expr{&&}, \expr{||}
	\item[parenthesizes] \expr{()}
	\item[calls and fields accesses] compiled as Haxe expressions
\end{description} 

When comparing two values with \expr{==} or \expr{!=} and when one of them can be \ic{NULL}, the SQL generator is using the \expr{<=>} SQL operator to ensure that \expr{NULL == NULL} returns true and \expr{NULL != NULL} returns false.

\paragraph{Additional search syntax}

It is also possible to use anonymous objects to match exact values for some fields (similar to previous SPOD but typed :

\begin{lstlisting} 
User.manager.search({id:1, name: "Nicolas"})
// same as :
User.manager.search($id == 1 && $name == "Nicolas")
// same as :
User.manager.search($id == 1 && {name: "Nicolas"})
\end{lstlisting} 

Use if-conditions to generate different SQL based on Haxe variables (it is not possible to use database fields in if-test) :

\begin{lstlisting} 
function listName( ?name : String ) {
    return User.manager.search($id < 10 && if (name == null) true else $name == name);
}
\end{lstlisting} 

\subsection{SQL functions}
\label{std-spod-sql}

\paragraph{SQL global functions}
The following SQL global functions are possible in search queries:

\begin{center}
\begin{tabular}{| l | l | 1 | 1 |}
	\hline
	\multicolumn{4}{|c|}{functions} \\ \hline
	Function & Haxe Type & SQL Type &  Description \\ \hline
	\expr{$now()} & \expr{SDateTime} & \ic{SQL NOW()} & returns the current datetime \\
	\expr{$curDate()} & \expr{SDate} & \ic{SQL CURDATE()} & returns the current date \\
	\expr{$date(v:SDateTime)} & \expr{SDate} & \ic{SQL DATE()} & returns the date part of the \expr{DateTime} \\
	\expr{$seconds(v:Float)} & \expr{SInterval} & \ic{SQL INTERVAL v SECOND} & returns the date interval in seconds  \\
	\expr{$minutes(v:Float)} & \expr{SInterval} & \ic{SQL INTERVAL v MINUTE} & returns the date interval in minutes \\
	\expr{$hours(v:Float)} & \expr{SInterval} & \ic{SQL INTERVAL v HOUR} & returns the date interval in hours \\
	\expr{$days(v:Float)} & \expr{SInterval} & \ic{SQL INTERVAL v DAY} & returns the date interval in days \\
	\expr{$months(v:Float)} & \expr{SInterval} & \ic{SQL INTERVAL v MONTH} & returns the date interval in months \\
	\expr{$years(v:Float)} & \expr{SInterval} & \ic{SQL INTERVAL v YEAR} & returns the date interval in years \\
\end{tabular}
\end{center}

\paragraph{SQL global operations}
The following SQL global operators are possible in search queries:

\begin{center}
\begin{tabular}{| l | l | 1 |}
	\hline
	\multicolumn{3}{|c|}{operators} \\ \hline
	Search operator & SQL operator &  Description \\ \hline
	\expr{stringA.like(stringB)} & SQL LIKE & find if \ic{stringB} is contained into \ic{stringA} \\
	\expr{$name in ["a", "b", "c"]}  &  & Pass any \ic{Iterable} to the in operator. An empty iterable will emit a false statement to prevent SQL errors when doing \ic{IN ()}. \\
\end{tabular}
\end{center}

\subsection{Search options}
\label{std-spod-search-options}

After the search query, specify some search options:

\begin{lstlisting} 
// retrieve the first 20 users ordered by ascending name
User.manager.search(true, {orderBy: name, limit: 20});
\end{lstlisting} 

The following options are supported:

\begin{itemize} 
	\item[\expr{orderBy}] specify one of several order database fields and use a minus operation in front of the field to sort in descending order. For instance \ic{\{orderBy: \[-name, id\]\}} will generate \ic{SQL ORDER BY name DESC, id}
	\item[\expr{limit}] specify which result range to obtain. Use Haxe variables and expressions in limit values, for instance \ic{\{ limit: \[pos, length\]\}}
	\item[\expr{forceIndex}] force this search to use the specific index. For example to force a two-fields index use \ic{\{ forceIndex: \[name, date\]\}}. The index name used in that case will be \ic{TableName_name_date}
\end{itemize} 

\subsection{SPOD Dynamic search}
\label{std-spod-dynamic-search}

To build own exact-values search criteria at runtime, use \expr{Manager.dynamicSearch} that will build the SQL query based on the values you pass it. Please note that you can get runtime errors if your object contain fields that are not in the database table.

\begin{lstlisting} 
var query = {name: "John", phoneNumber: "+818123456"};
var users = User.manager.dynamicSearch(query);
\end{lstlisting} 

\subsection{SPOD Relation search}
\label{std-spod-relation-search}

Search a given \tref{relation}{std-spod-relations} by using either the relation key or the relation property.

\begin{lstlisting} 
var user = User.manager.get(1);
var groups = UserGroup.manager.search($uid == user.id);

// same as:
var groups = UserGroup.manager.search($user == user);
\end{lstlisting} 
The second case is more strictly typed since it does not only check that the key have the same type, and it also safer because it will use null id if the user value is null at runtime.

\subsection{SPOD Relations}
\label{std-spod-relations}

Declare relations between database classes by using the \ic{@:relation} metadata.

\begin{lstlisting} 
class User extends sys.db.Object {
    public var id:SId;
    // ....
}
class Group extends sys.db.Object {
   public var id:SId;
   // ...
}

@:id(gid, uid)
class UserGroup extends sys.db.Object {
    @:relation(uid) public var user:User;
    @:relation(gid) public var group:Group;
}
\end{lstlisting} 

The first time the user field from an UserGroup instance gets read by SPOD Macros, SPOD will fetch the User instance corresponding to the current uid value and cache it. When the user field gets set, it will modify the uid value as the same time.

\paragraph{Locking}
When using transactions, the default behavior for relations is that they are not locked. Lock the row (\ic{SQL SELECT...FOR UPDATE}) by adding the lock keyword after the relation key:

\begin{lstlisting} 
@:relation(uid, lock) public var user:User;
\end{lstlisting} 

\paragraph{Cascading}

Relations can be strongly enforced by using \ic{CONSTRAINT}/\ic{FOREIGN KEY} with MySQL/InnoDB. This way when an User instance is deleted, all the corresponding UserGroup for the given user will be deleted as well. However if the relation field can be nullable, the value will be set to \ic{NULL}. To enforce cascading for nullable-field relations, add the cascade keyword after the relation key.

\begin{lstlisting} 
@:relation(uid,cascade) var user:Null<User>;
\end{lstlisting} 


\subsection{SPOD Serialized Data}
\label{std-spod-serialized-data}

\since{Haxe 2.11}

In order to store arbitrary serialized data in a SPOD object, use the \expr{SData} type. For example:

\begin{lstlisting} 
import sys.db.Types
enum PhoneKind {
    AtHome;
    AtWork;
    Mobile;
}
class User extends sys.db.Object {
    public var id : SId;
    ...
    public var phones : SData<Array<{kind:PhoneKind, number:String}>>;
}
\end{lstlisting} 

\begin{itemize}
	\item When the phones field is accessed for reading (the first time only), it is unserialized. By default the data is stored as an Haxe-serialized String, but is possible to override the \expr{doSerialize} and \expr{doUnserialize} methods of the Manager to have a specific serialization for a specific table or field.
	\item When the phones field has been either read or written, a flag will be set to remember that potential changes were made.
	\item When the SPOD object is either inserted or updated, the modified data is serialized and eventually sent to the database if some actual change have been done.
\end{itemize}

As a consequence, pushing data into the phones \type{Array} or directly modifying the phone number will be noticed by the SPOD engine.

The SQL data type for \expr{SData} is a binary blob, in order to allow any kind of serialization (text or binary), so the actual runtime value of the phones field is a Bytes. It will however only be accessible by reflection, since SPOD is changing the phones field into a property.

\paragraph{Accessing the SPOD Infos}
Get the database schema by calling the \expr{Manager.dbInfos()} method. It will return a \type{sys.db.SpodInfos} structure.
